Quantifying the importance of a tree prior in Bayesian phylogenetics

### 2019-04-15 article.tex
Tree from unknown tree prior here

The second step (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a)
selects inference models from a set of experiments. 
We define an experiment $X$ as the combination of an inference model $I$
and the conditions $C$ to actually use it in the inference step.
There are two main setups of condition $C$: either the attached
inference model $I$ is the generative model and should always be selected,
or the attached inference model is a candidate model, that will be
be selected if it is the best model \giovanni{then why do you need the "run\_if" option?}.
When setting $C$ to be attached to the generative model,
it is implied that attached inference model $I$ uses the 
same site and clock model used in the alignment simulation,
and $I$ uses the same tree prior as used in the creation of the given phylogeny. 
A 'best' model is defined as the inference model with
the highest evidence (a.k.a. marginal likelihood) for that inference model, 
given the alignment simulated in the previous step.
The evidence of an inference model is estimated by nested 
sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package.
Scripted use of BEAST2 packages is only possible under under Linux and Mac.
Windows systems can do the model comparison for shorter DNA sequences
using the web interface of \verb;mcbette; [\cite{mcbette}].

@@

The fourth step quantifies the inference error made.
It creates an error distribution, by measuring the difference
between the true tree and each of the posterior 
trees (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
The default way to quantify the difference between two phylogenies
is the nLTT statistic (\cite{janzen2015approximate}), but any 
user-defined error statistic can be used.
\giovanni{How is the MCMC topic connected to what previously said?} An MCMC needs to 'warm up', as initially it will sample
an unrepresentative part of state space. This phase, called 
the burn-in, is commonly set up to be a small part of the entire run, 
after which it is discarded. By default, \verb;pirouette; 
uses a cut-off value of 10\%.

@@

\begin{lstlisting}[
  language=R, 
  floatplacement=ht, 
  frame=single, 
  label = {lst:unknown_phylogeny},
  caption = A phylogeny generated by an unknown diversification model.
]
phylogeny  <- ape::read.tree(
  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
)
\end{lstlisting}
\begin{figure}[ht]
  \includegraphics[width=\textwidth]{example_2/tree_unknown.png}
  \caption{The tree derived from an unknown diversification process, 
    as created by listing~\ref{lst:unknown_phylogeny}.
  }
\end{figure}

@@

We choose again to use the default error measurement setup,
as shown in listing~\ref{lst:create_error_measure_params},
which is the default in \verb;create_pir_params;, so we'll
leave it out for now.

>>>>

The second step (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a)
selects inference models from a set of experiments. 
We define an experiment $X$ as the combination of an inference model $I$
and the conditions $C$ to actually use it in the inference step.
There are multiple setups of condition $C$: 
either the attached inference model $I$ is the generative model and should always be selected,
or the attached inference model is a candidate model, that will be
be selected if it is the best model, or will be selected always (regardless
of it being the best). 
When setting $C$ to be attached to the generative model,
it is implied that attached inference model $I$ uses the 
same site and clock model used in the alignment simulation,
and $I$ uses the same tree prior as used in the creation of the given phylogeny. 
A 'best' model is defined as the inference model with
the highest evidence (a.k.a. marginal likelihood) for that inference model, 
given the alignment simulated in the previous step.
The evidence of an inference model is estimated by nested 
sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package.
Scripted use of BEAST2 packages is only possible under under Linux and Mac.
Windows systems can do the model comparison for shorter DNA sequences
using the web interface of \verb;mcbette; [\cite{mcbette}].

@@

The fourth step quantifies the inference error made.
It creates an error distribution, by measuring the difference
between the true tree and each of the posterior 
trees (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
The default way to quantify the difference between two phylogenies
is the nLTT statistic (\cite{janzen2015approximate}), but any 
user-defined error statistic can be used.
The error statistic is not applied on all posterior phylogenies 
\giovanni{How is the MCMC topic connected to what previously said?} 
XXXXXXXXXXXXXXXXXXXXXXXXXXXX33
An MCMC needs to 'warm up', as initially it will sample
an unrepresentative part of state space. This phase, called 
the burn-in, is commonly set up to be a small part of the entire run, 
after which it is discarded. By default, \verb;pirouette; 
uses a cut-off value of 10\%.

@@

\begin{lstlisting}[
  language=R, 
  floatplacement=ht, 
  frame=single, 
  label = {lst:unknown_phylogeny},
  caption = A phylogeny generated by an unknown diversification model.
]
phylogeny  <- ape::read.tree(
  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
)
\end{lstlisting}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_2/tree_unknown.png}
  \caption{The tree derived from an unknown diversification process, 
    as created by listing~\ref{lst:unknown_phylogeny}.
  }
\end{figure}

@@

We choose again to use the default error measurement setup,
as shown in listing~\ref{lst:create_error_measure_params}.
